---
title: "Differential Gene Expression Analysis of Breast Cancer"
author: "Raksha Devi"
date: "`r Sys.Date()`"
output: html_document
---

## Introduction

This project analyzes real breast cancer data to compare
benign and malignant tissue samples using R Markdown.

---

## Load Required Libraries

```{r}
library(tidyverse)
library(ggplot2)

#load Dataset=>
data <- read.csv("data/gene_expression.csv")
head(data)

#Data Cleaning=>
data_clean <- na.omit(data)

```

## Biological Meaning of Features

The dataset contains cellular characteristics extracted from breast tissue biopsies.

- Cl.thickness: Thickness of cell clusters
- Cell.size: Uniformity of cell size
- Cell.shape: Uniformity of cell shape
- Marg.adhesion: Cell adhesion properties
- Epith.c.size: Epithelial cell size
- Bare.nuclei: Presence of bare nuclei
- Bl.cromatin: Chromatin texture
- Normal.nucleoli: Nucleoli appearance
- Mitoses: Cell division activity
- Class: 0 = Benign, 1 = Malignant

## Statistical Test: cell Thickness=>
```{r}
t_test_result <- t.test(
  Cl.thickness~Class,
  data = data_clean
)
t_test_result
```
## Distribution of samples=>
```{r}
table(data_clean$Class)
```

# Boxplot=>
We now **visualize** how cell thickness differs between classes.


## cell Thickness Comparsion=>
```{r}
ggplot(data_clean, aes(x = factor(Class), y = Cl.thickness))+
  geom_boxplot(fill= "steelblue")+
  labs(
    title = "cell Thickness in Benign vs Malignant Samples",
    x = "Class(0=benign , 1 = Malignant )",
    y = "Cell Thickness"
  )+
  theme_minimal()
```
## Save Cell Thickness Boxplot=>
```{r}
ggsave(
  filename = "results/cell_thickness_boxplot.png",
  width = 6,
  height = 5
)
```
## Multi-feature Conparsion
This compares ** all cellular features at once**.
```{r}
data_long <- data_clean %>%
  pivot_longer(
    cols = Cl.thickness:Mitoses,
    names_to = "Feature",
    values_to = "Value"
    
  )
ggplot(data_long,aes(x = factor(Class), y = Value))+
  geom_boxplot()+
  facet_wrap(~Feature, scales = "free")+
  labs(
    title = "Comparison of Cellular Features Between Classes",
    x = "Class (0 = Benign , 1 = Malignant)",
    y = "Feature Value"
  )+
  theme_minimal()
```
## Save Multi-Feature Plot=>
```{r}
ggsave(
  filename ="results/multifeature_boxplot.png",
  width= 10,
  height = 6
)



















